### 2015-06-30 Projects/DelCo_2012/Report/DelCo_2012.tex
Delco_2012: minor report edit
DelCo2012: update report

The structures were instrumented for temperature, gas velocity, and heat flux measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single thermocouples were installed in conjunction with the bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples, with a 1.0~mm (0.04~in) nominal diameter. The thermocouple wire was protected with an 3.2~mm (0.125~n) diameter inconnel sheath. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to isolate radiative heat transfer.

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/DelCoSingleStoryInstrumentationDimensions}
	\caption{Photograph of the Instrumentation Dimensions in the Fire Suppression Experiments.}
	\label{fig:Fire_Suppression_Instrumentation_Dimensions}
\end{figure}

@@

There are different components of uncertainty in the length, mass, temperature, heat flux, gas concentration, differential pressure, gas velocity and heat release rate reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means  [ ]. Type B analysis of systematic uncertainties involves estimating the upper (+ a) and lower (- a) limits for the quantity in question such that the probability that the value would be in the interval (± a) is essentially 100~\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95~\% confidence interval (2σ).  For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

Each length measurement was taken carefully. Length measurements such as the room dimensions, instrumentation array locations and fire apparatus (for example nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which is has an accuracy of ± 6.0 mm (0.25) over a range of 0.61 m (2.00 ft) to 15.3 m (50.0 ft) [ ] .  However, conditions affecting the measurement, such as levelness of the device, yields an estimated uncertainty of ± 0.5~\% for measurements in the 2.0 m (6.6 ft) to 10.0 m (32.8 ft) range.  Steel measuring tapes with a resolution of  ± 0.5 mm (0.02 in) were used to locate individual sensors within a measurement array and to measure and position the furniture. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of ±1.6 mm (0.06 in) for 9.1 m (30 ft) tapes and ±6.4 mm (0.25 in) for 30.5 m (100 ft) tapes [ ] .  Some issues, such as “soft” edges on the upholstered furniture, result in an estimated total expanded uncertainty of  ± 1.0~\%.

The load cell used to weigh the fuels prior to the experiments had a range of 0 kg (0 lbs) to 200 kg (440lbs) with a resolution of a 0.05 kg (0.11 lb) and a calibration uncertainty within 1~\% [] . The expanded uncertainty is estimated to be less than  ± 5 %. 
The standard uncertainty in temperature of the thermocouple wire itself is  ± 2.2 °C at 277 °C and increases to ± 9.5 °C at 871 °C as determined by the wire manufacturer [ ] .  The variation of the temperature in the environment surrounding the thermocouple is known to be much greater than that of the wire uncertainty [ ,  ] .  Small diameter thermocouples were used to limit the impact of radiative heating and cooling.  The estimated total expanded uncertainty for temperature in these experiments is ± 15 %.
In this study, total heat flux measurements were made with water-cooled Schimidt-Bolter gauges.  The manufacturer reports a ± 3~\% calibration expanded uncertainty for these devices [ ] . Results from an international study on total heat flux gauge calibration and response demonstrated that the uncertainty of a Schmidt-Boelter gauge is typically ± 8~\% [] .

The gas measurement instruments and sampling system used in this series of experiments have been demonstrated an expanded (k = 2) relative uncertainty of ± 1~\% when compared with span gas volume fractions [].   Given the non-uniformities and movement of the fire gas environment and the limited set of sampling points in these experiments an estimated uncertainty of ± 12~\% is being applied to the results [ ].

Differential pressure reading uncertainty components were derived from pressure transducer
instrument specifications and previous experience with pressure transducers. The transducers
were factory calibrated and the zero and span of each was checked in the laboratory prior to the
experiments yielding an accuracy of ± 1~\% [ ] . The total expanded uncertainty was estimated at 10~\%.

Bi-directional probes and single thermocouples were used to measure the velocity.  The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above.   Bare-bead Type K thermocouple are co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from ± 0.14 to ± 0.22 for bi-directional probes of similar design[] .  The total expanded uncertainty for gas velocity in these experiments w estimated to be  ± 18 %.   
Water Flowrate
Water flowrate was measured with a pressure and flow meter combination shown in Figure x.x. The meter consists of a section of 6.35 cm (2.5 inch) cast aluminum pipe with a 0 – 4.1 MPa (0 - 600 psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800 lpm (1250 gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate.  The manufacturer reports a ± 5~\% calibration expanded uncertainty for the flow sensor and ± 3~\%  for the pressure sensor [ ] . The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flowrate. The total expanded uncertainty was estimated at ± 10 %.

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/DelCoSingleStoryWoodFuelLoad}
	\caption{Wood Fuel Load Dimensions.}
	\label{fig:Wood_Fuel_Load_Dimensions}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/DelCoSingleStoryFurnitureFuelLoad}
	\caption{Furniture Fuel Load Dimensions.}
	\label{fig:Furniture_Fuel_Load_Dimensions}
\end{figure}

>>>>

The structures were instrumented for temperature, gas velocity, and heat flux measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single thermocouples were installed in conjunction with the bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples, with a 1.0~mm (0.04~in) nominal diameter. The thermocouple wire was protected with an 3.2~mm (0.125~n) diameter inconnel sheath. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer.

@@

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Pictures/DelCoSingleStoryInstrumentationDimensions}
	\caption{Photograph of the Instrumentation Dimensions in the Fire Suppression Experiments.}
	\label{fig:Fire_Suppression_Instrumentation_Dimensions}
\end{figure}

@@

There are different components of uncertainty in the length, mass, temperature, heat flux, gas concentration, differential pressure, gas velocity and heat release rate reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+ a) and lower (- a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$ a) is essentially 100~\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95~\% confidence interval (2$\sigma$).  For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

Each length measurement was taken carefully. Length measurements such as the room dimensions, instrumentation array locations and fire apparatus (for example nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which is has an accuracy of $\pm$ 6.0 mm (0.25) over a range of 0.61 m (2.00 ft) to 15.3 m (50.0 ft)~\cite{StanleyTools}.  However, conditions affecting the measurement, such as levelness of the device, yields an estimated uncertainty of $\pm$ 0.5~\% for measurements in the 2.0~m (6.6~ft) to 10.0~m (32.8~ft) range.  Steel measuring tapes with a resolution of  $\pm$ 0.5~mm (0.02~in) were used to locate individual sensors within a measurement array and to measure and position the furniture. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of $\pm$1.6~mm (0.06~in) for 9.1~m (30~ft) tapes and $\pm$ 6.4~mm (0.25~in) for 30.5~m (100~ft) tapes~\cite{Butcher:2012}. Some issues, such as ``soft'' edges on the upholstered furniture, result in an estimated total expanded uncertainty of $\pm$ 1.0~\%.

The load cell used to weigh the fuels prior to the experiments had a range of 0~kg (0~lb) to 200~kg (440~lb) with a resolution of a 0.05~kg (0.11~lb) and a calibration uncertainty within 1~\%~\cite{Ohaus:2000}. The expanded uncertainty is estimated to be less than $\pm$ 5~\%.

The standard uncertainty in temperature of the thermocouple wire itself is $\pm$ 2.2 $^{\circ}$C at 277 $^{\circ}$C and increases to $\pm$ 9.5~$^{\circ}$C at 871~$^{\circ}$C as determined by the wire manufacturer\cite{Omega:2004}. The variation of the temperature in the environment surrounding the thermocouple is known to be much greater than that of the wire uncertainty~\cite{Blevins:1999,Pitts:2003}. Small diameter thermocouples were used to limit the impact of radiative heating and cooling. The estimated total expanded uncertainty for temperature in these experiments is $\pm$ 15~\%.

In this study, total heat flux measurements were made with water-cooled Schimidt-Bolter gauges. The manufacturer reports a $\pm$ 3~\% calibration expanded uncertainty for these devices~\cite{Medtherm:2003}. Results from an international study on total heat flux gauge calibration and response demonstrated that the uncertainty of a Schmidt-Boelter gauge is typically $\pm$ 8~\%~\cite{Pitts:2006}.

The gas measurement instruments and sampling system used in this series of experiments have been demonstrated an expanded (k = 2) relative uncertainty of $\pm$ 1~\% when compared with span gas volume fractions~\cite{Bundy:2007}. Given the non-uniformities and movement of the fire gas environment and the limited set of sampling points in these experiments an estimated uncertainty of $\pm$ 12~\% is being applied to the results~\cite{Lock:1}.

Differential pressure reading uncertainty components were derived from pressure transducer instrument specifications and previous experience with pressure transducers. The transducers were factory calibrated and the zero and span of each was checked in the laboratory prior to the experiments yielding an accuracy of $\pm$ 1~\%~\cite{Setra:2002}. The total expanded uncertainty was estimated at 10~\%.

Bi-directional probes and single thermocouples were used to measure the velocity.  The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above. Bare-bead Type K thermocouple are co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from $\pm$ 0.14 to $\pm$ 0.22 for bi-directional probes of similar design~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments w estimated to be  $\pm$ 18~\%.

Water flowrate was measured with a pressure and flow meter combination. The meter consists of a section of 6.35~cm (2.5~inch) cast aluminum pipe with a 0 - 4.1~MPa (0 - 600~psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800~lpm (1250~gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate.  The manufacturer reports a $\pm$ 5~\% calibration expanded uncertainty for the flow sensor and $\pm$ 3~\%  for the pressure sensor~\cite{Akron:2009}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flowrate. The total expanded uncertainty was estimated at $\pm$ 10 %.

@@

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Pictures/DelCoSingleStoryWoodFuelLoad}
	\caption{Wood Fuel Load Dimensions.}
	\label{fig:Wood_Fuel_Load_Dimensions}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Pictures/DelCoSingleStoryFurnitureFuelLoad}
	\caption{Furniture Fuel Load Dimensions.}
	\label{fig:Furniture_Fuel_Load_Dimensions}
\end{figure}
